%
%  VISTA Cookbook;  Master File
%
%  Began:  1988 August 30
%
%

\documentstyle {manual}
%\input sys$user:[rick.thesis]thesismacros.pogge
\input thesismacros.pogge

\newenvironment{command}{\begin{center}
\begin{list}{\tt GO:}{\setlength{\rightmargin}%
{\leftmargin}}\tt\singlespace }{\end{list}\end{center}}

\def \comm#1{{\tt #1\/}}
\def \hitkey#1{$\langle${\tt {#1}}$\rangle$\/ }

\begin{document}
\pagestyle{manual}
\singlespace

\pagenumbering{roman}

\clearpage
\typeout{Title Page}

\pagestyle{empty}

\begin{center}

   University of California\\
   \vbox to 0.15in {}
   Lick Observatory Technical Reports\\
   \vbox to 0.15in {}
   No. 50\\
   \vbox to 1.5in {}
   {\Huge The VISTA Cookbook}\\
   \vbox to 1.0in {}
   Richard W. Pogge\\
   Robert W. Goodrich\\
   Sylvain Veilleux\\
   \vbox to 0.25in {}
   Santa Cruz, California\\
   \vbox to 0.15in {}
   August 1988\\
\end{center}

\vfill
\eject

%

\singlespace
\clearpage
\typeout{Copyright Page}
\pagestyle{empty}

\vbox to 1.5in {}

\noindent{{\small\rm Copyright~\copyright~1988 by U.C. Regents}}

\vbox to 0.2in {}

\noindent{{\small\rm This document is intended to be used in conjuction with
the VISTA Image Processing Package distributed free of charge by the Lick
Observatory, University of California.  Extensive copying of this manual, in
whole or in part, is allowable only for scholarly purposes, consistent with
``fair use'' as described in the U.S. Copyright Law.  Copies may be obtained
from the Publications Office of the Lick Observatory, University of
California, Santa Cruz, CA 95064.}}

\vbox to 0.2in {}

\noindent{{\small\rm Composed in Times Roman on the Lick Observatory VAX
11/780 computer using Leslie Lamport's \LaTeX, and printed on an Imagen 8/300
laser printer.  Figures were composed in PostScript and printed on an Apple
LaserWriter II-NTX connected to a Sun 4/280.}}

\vfill

\noindent{{\bf On the Cover:\ } CCD image of the Crab Nebula Supernova Remnant
in the light of the [\sii]~\waves{6716,~6731} emission-lines.  Taken with the
TI 500$\times$500 CCD and the focal reducing camera on the 1-meter Anna Nickel
Telescope at Mt. Hamilton.  The image was processed and prepared for display
by Rick Pogge using VISTA.}

\eject

{\singlespace
\pagestyle{manual}
\tableofcontents}

\vfill
\eject

\input{cook_intro}

\input{cook_tutor}

\input{cook_basic}

\input{cook_2d}

\input{cook_spec}

\input{cook_ham}

\input{cook_dirt}

\appendix

\input{cook_appa}

\input{cook_appb}

\input{cook_appc}

\end{document}
\end
